\chapter{Literature Review on CVRP exact methods}
\label{sec:intro-literature-review}

This chapter lists some significant contributions regarding exact approaches
in the CVRP and VRPTW domains.

\textcite{dantzig1959} were
the pioneers of the contemporary vehicle routing problem two
introduced the CVRP problem to the masses.
A few years later, \textcite{clarke1964} proposes
an effective greedy heuristic algorithm for tackling the CVRP.
The presentation of this new problem led to a whole new branch
of scientific research within the operations research community.
In the last 60 years, a huge number of contributors in
the operations research community has studied and proposed many
mathematical models and algorithms, both exact and (meta-)heuristic,
for solving the VRP.

Until roughly the late 1980s, exact approaches to the CVRP were
based on tree-search algorithms employing branch-and-bound schemes
(see \cite{pierce1969, christofides1969, christofides1981exact, laporte1986}),
occasionally employing Lagrangian duality relaxation (see \cite{fisher1994, miller1995}),
or additive bounding procedures (see \cite{fischetti1994a, hadjiconstantinou1995}).

Some authors applied branch-and-cut (BAC) algorithms  to the VRP from the 1980s to the early 2000s.
\textcite{laporte1983, laporte1985} proposed the two-index arc flow formulation
and a branch-and-cut algorithm for optimally solving the CVRP.
\Textcite{augerat1995approche}, later proposed a branch-and-cut scheme
integrating additional valid inequalities.
A few years later, \textcite{lysgaard2004}
proposed a branch-and-cut framework and
a set of improved separation procedures for the valid CVRP inequalities that were known at the time.
Other notable BAC-based contributions
can be found in \textcite{araqueg1994, augerat1995, achuthan1996, blasum2000, ralphs2003, achuthan2003, baldacci2004}.
\citeauthor{baldacci2004} was able to solve a 135 customers-sized routing problem
from the F test-set  \textcite{fisher1994}
for the first time.

Despite the promising results, nonetheless,
the branch-and-cut contributions of the time demonstrated
that CVRP instances with even a few customers (less than 100)
proved exceedingly challenging to solve exactly.

\Textcite{desrosiers1984, agarwal1989setpartitioningbased}
are a few very early attempts at applying the \textit{Set Partitioning} (SP) formulation
and a \textit{column generation} scheme to routing problems.
The \textit{Set Partitioning} formulation and the \textit{column generation}
will be  presented in more details in \cref{sec:set-partitioning-formulation, sec:column-generation-and-pricing-problem}
respectively.
These authors demonstrated that the column generation scheme proved very satisfactory
for solving the VRPTW under tight constraints
but did not achieve remarkable results for the CVRP at the time.
As a result, until the early 2000s,
branch-and-cut approaches were deemed the best approach for solving the CVRP.

In \citeyear{desrochers1992}, \citeauthor{desrochers1992}
proposed a dynamic programming-based label-correcting algorithm for the SPPRC,
and devised it within a \textit{branch-and-price} (BAP) scheme to tackle the VRPTW.
This labeling algorithm is still a key component of modern BPC schemes
for efficiently solving the \textit{pricing problem}.
\textit{Branch-and-price} schemes and the \textit{pricing-problem}
will be both discussed in the dedicated \cref{sec:branch-and-price}.
The 2-paths inequalities were introduced \textcite{kohl1999},
which were later generalized in \textcite{desaulniers2008}.
\Textcite{kohl1999} was one of the first works to attempt
to integrate cutting planes within a BAP framework for solving the CVRP.

\Textcite{fukasawa2006} achieved one of the outstanding breakthroughs in column generation and VRP research.
\Citeauthor{fukasawa2006} were the first authors to propose a BPC algorithm
that incorporated a powerful robust-cuts-based cutting-plane approach normally found in BAC algorithms.
\citeauthor{fukasawa2006} integrated the cutting-planes proposed in \cite{lysgaard2004},
and the column generation considered only $q$-routes with $k$-cycles elimination.
See \cref{sec:branch0-and-price, sec:solving-the-pricing-problem} for a thorough introduction.
\textcite{fukasawa2006} was able to solve many CVRP instances for up to $100$ customers approximately.
From the work of \citeauthor{fukasawa2006} until now,
the column generation and the label-correcting algorithm
have stood as the de facto procedures of modern and efficient VRP solvers.

\Textcite{baldacci2008} later proposed a slightly different approach compared to \textcite{fukasawa2006}.
They employ a column generation scheme considering elementary routes
with additional cutting planes.
Furthermore, they used a bounding procedure to improve the dual bound
and reduce the number of routes to consider.
They turned off branching and used \textit{route enumeration}
at the root node to generate a comprehensive MIP model to be solved directly.
Refer to \cref{sec:branching-and-cut-generation-within-bap-frameworks} for additional details
regarding \textit{route enumeration}.
\textcite{pessoa2008} improves over the work of \textcite{fukasawa2006}
by integrating the works of \textcite{fukasawa2006, baldacci2008}
through hybridization between traditional branching and enumeration.
Instead, \textcite{contardo2011} takes a different approach.
It considers $q$-routes with 2-cycle eliminations and employs non-robust k-CECs cuts
to enforce elementarity.

The ng-routes, first proposed in \textcite{baldacci2011},
are a very effective partial (soft) elementarity relaxation
commonly found in contemporary efficient BPCs for the VRP.
The ng-routes relaxation significantly reduces the pricing problem's complexity
while producing reasonable dual-bounds.
\textcite{ropke2012} revisits the work \citeauthor{fukasawa2006}
and applies the ng-routes relaxation and strong branching schemes.
Refer to \cref{sec:solving-the-pricing-problem} for a thorough introduction to the ng-routes relaxation.

\Textcite{contardo2014} outperforms \textcite{contardo2011}
by proposing: an effective enumeration scheme with
$q$-routes with 2-cycle elimination, ng-routes relaxation with ng-sets of size $8$,
non-robust $k$-CECs cuts to avoid larger cycles
and the \textit{decremental state-space relaxation} technique.
They were able to solve CVRP instances made up of $151$ customers using this approach.
The \textit{decremental state-space relaxation} is extended to handle ng-sets in \textcite{martinelli2014},
where the size of the ng-sets size is dynamically enlarged to seek better dual bounds.
For more information on the \textit{decremental state-space relaxation},
see \cref{sec:solving-the-pricing-problem}.

\Textcite{pecin2017improved} proposed another modern BCP algorithm for solving the CVRP.
They solved $200$-customers-CVRP instances and with some successes in solving some $360$-customers instances.
Their work featured the following modern components:
(i) bidirectional label-setting algorithm for solving the pricing problem,
(ii) ng-routes relaxation,
(iii) non-robust lm-SRCs cuts,
(iv) modern dual smoothing
and (v) route enumeration.
\Textcite{pecin2017new} extends the work of \textcite{pecin2017improved} to the VRPTW.
On the day of writing, \textcite{pessoa2018enhanced, pessoa2020generic}
are the current state-of-the-art BCP framework for solving the CVRP
based on column generation with ng-routes relaxation.
\Textcite{pessoa2020generic} was able to optimally solve six opened CVRP instances
from the X test-set \parencite{uchoa2017} of up to $548$ customers.

Due to significant advances in exact BPC algorithms over the last decades,
today,
we are able to optimally solve CVRPs with more than $300$ customers \parencite{costa2019}.

Once we introduce the pricing sub-problem induced by BAP/BPC frameworks
in \cref{sec:branch-and-price},
we will discuss additional contributions regarding this problem domain,
see \cref{sec:solving-the-pricing-problem}.
For a literature review on branch-and-cut approaches for pricing, see later discussion
in \cref{sec:bac-approaches-for-the-pricing-problem}.
