\chapter{The Pricing Problem}
\label{sec:the-pricing-problem}

As it was already discussed in \cref{sec:intro-branch-and-price},
the Elementary Shortest Path Problem with Resource Constraints (ESPPRC)
appears as the pricing sub-problem in column generation schemes
for vehicle routing problems.
Therefore, its study is of fundamental importance to
engineer efficient branch-and-price frameworks.
We refer the reader to \cref{sec:intro-solving-the-pricing-problem}
for a literature review on the importance and methodologies
for solving the pricing problem.

While the ESPPRC can be seldom studied as a standalone problem,
in this work we are mainly interested at it from the perspective
of pricing for the vehicle routing problem.
The ESPPRC asks for the determination of a shortest path (or route)
between a starting and destination vertex over a directed weighted network
subject to additional resource restrictions.
The general version of the ESPPRC allows for multiple resources definition
with different semantics, for example: time, capacity, gasoline, etc.
The resources are usually monotically increasing quantities which
increase as the route traverses new edges or vertices \parencite{irnich2005, irnich2007resource}.
Resource consumption bounds can be given at different level of scopes which
can be classified as: at the global scope, at the vertex level scope, or at the arc level scope.

The Elementary Shortest Path Problem with Capacity Constraints (ESPPCC) is a special
case of ESPPRC characterized by a single resource, the amount of goods,
which is bounded by a global quantity $Q \in \R$, the vehicle capacity.
The ESPPCC can be used to model the pricing sub-problem in column
generation schemes for the Capacitated Vehicle Routing Problem (CVRP).

In this work, we will mostly focus our attention to the ESPPCC pricing sub-problem.
It is important to acknowledge however, that the
more general ESPPRC is required to model more complex scenarios,
such as for example, the pricing sub-problem induced
from the Vehicle Routing Problem with Time Windows (VRPTW).

The Capacitated Profitable Tour Problem (CPTP) is a combinatorial optimization problem
which, alike the ESPPCC,
can be effectively used to model the pricing sub-problem
induced from the CVRP.
The CPTP asks for a minimum cost circuit subject to capacity restrictions,
and can be seen as a special case of ESPPCC where the starting
and destination vertices coincide.

\medskip

We will formally introduce, discuss and provide an IP formulation
for the ESPPCC in \cref{sec:the-elementary-shortest-path-problem-with-capacity-constraints}.
In \cref{sec:the-capacitated-profitable-tour-problem}, instead,
we will introduce the CPTP along with an IP formulation and additional valid inequalities.
We will study the basic version of these two problems, namely
non-robust inequalities altering the structure of the problem, are not considered.
However, before diving directly to the discussion of these two problems,
we will first dedicate the next section (\labelcref{sec:the-pricing-problem-literature-review})
to review the few past contributions regarding
the usage of branch-and-cut approaches for pricing.

\subsection{Literature Review on Branch-and-Cut approaches applied to the Pricing Problem}
\label{sec:the-pricing-problem-literature-review}

Label-correcting dynamic programming approaches
dominate as the main approach to solve the (E)SPPRC in the context of pricing,
see \cite{desrochers1992, feillet2004, righini2004, righini2006, boland2006, righini2008, pugliese2010, baldacci2011, lozano2013, lozano2016, sadykov2021bucket}.
The label correcting algorithm is a natural extension of the Bellman–Ford algorithm \parencite{bellman1958, fordjr1956},
a common algorithm present in the "toolbelt" of all operations research practitioners,
which computes shortest paths from a source vertex to all other vertices,
by taking also in consideration additional path constraints.

To the best of our knowledge,
branch-and-cut approaches for tackling the (E)SPPRC have been barely covered.
The few branch-and-cut contributions that we are aware of are:
\textcite{jepsen2008branchandcut} for the ESPPCC,
\textcite{jepsen2011,jepsen2014} to the CPTP,
\textcite{taccari2016, drexl2014} for the related non-resource constrained ESPP,
\textcite{horvath2016} for the (non-elementary) RCSP.
Branch-and-cut solutions for the general ESPPRC version have not been covered.
This situation can probably be attributed to the
inherent difficulties associated in modeling
resource constraints with non-global bounds \parencite{jepsen2008branchandcut},
such as the time window constraints .
First, non-global bounds on the resources make mandatory the usage of
a directed network based IP model,
which is twice times larger than an undirected equivalent solution.
Second, the non-globally bounded resource constraints may be only modeled as
big-M constraints, which are known to be inherently computationally unstable \parencite{jepsen2008branchandcut}.

\medskip

\textcite{jepsen2008branchandcut} provides an IP mathematical model
for the ESPPCC over an undirected network, along with additional valid inequalities.
In their work they empirically evaluate the effectiveness of their
branch-and-cut framework in the context of pricing for the CVRP.

\textcite{jepsen2014} study the CPTP problem in the context of pricing for the CVRP.
Their work originate from the initial efforts of \textcite{jepsen2008branchandcut}.
The work of \citeauthor{jepsen2014} provides a first tutorial/survey/framework
and foundational theory for further development of the CPTP as, either
an independent problem or in the context of pricing.
They focus on the development of a model, additional valid inequalities and
a branch-and-cut algorithm for solving it.
They readjust many inequalities from related problems
and discuss their efficient separation.

\textcite{jepsen2014} study the competitiveness of their approach
along with the usefulness of the valid inequalities through a computational study.
Their results show that the CPTP based branch-and-cut framework
was not competitive w.r.t. to dynamic programming
in solving the pricing induced from the column generation for CVRPs.
However, their results showed that the branch-and-cut algorithm
was able to solve some bigger ESPPCC/CPTP instances where the
dynamic programming solution suffered.
Their results show that:
\begin{enumerate}
	\setlength{\itemsep}{0pt}
	\setlength{\parskip}{0pt}

	\item \label{itm:cptp-branch-and-cut-better-p1} The branch-and-cut
	      algorithm seemed to help in solving some standalone CPTP instances characterized
	      by a huge number of customers (approximately more than 200),
	      where the labeling algorithm ran out of available computation time.
	\item \label{itm:cptp-branch-and-cut-better-p2} The branch-and-cut algorithm
	      performed much better w.r.t. the labeling algorithm when the
	      weights of the associated ESPPCC network were highly negative.
	\item \label{itm:cptp-branch-and-cut-better-p3} The branch-and-cut framework had a tendency
	      to perform worse when the optimal objective value of the problem was very close to the zero
	      threshold.
\end{enumerate}
The result \labelcref{itm:cptp-branch-and-cut-better-p1} is obtained for standalone
CPTP solutions and thus may not be relevant in the context of pricing.
Result \labelcref{itm:cptp-branch-and-cut-better-p2} is of little-interest
in the context of pricing
since highly negative weights network are usually solved through heuristics regardless.
The result \labelcref{itm:cptp-branch-and-cut-better-p3}, which is not in favor for a
branch-and-cut based pricing solution,
can probably be attributed to the fact that labeling algorithms
are usually optimized to cut off huge portions of the solution space when
the objective function has optimal values very close to the zero threshold.

Despite their contribution needs to be revisited given the modern advances on both sides,
their results proved that a branch-and-cut approach could complement
the dynamic programming algorithm for pricing.
In conclusion, there might value revisiting and
expanding the research branch-and-cut based pricers based on the CPTP formulation.

\section{The Elementary Shortest Path Problem with Capacity Constraints}
\label{sec:the-elementary-shortest-path-problem-with-capacity-constraints}

This section is dedicated to the discussion of the
Elementary Shortest Path Problem with Capacity Constraints (ESPPCC)
in the context of pricing for the CVRP.
While we have already glimpsed at the general ESPPRC case of this problem in
\cref{sec:column-generation-and-pricing-problem,sec:intro-solving-the-pricing-problem},
we've never provided a formal description of the mathematical model.
In this section we consider the basic version of the ESPPCC,
i.e. non-robust inequalities at the RMP level are not considered.
In case non-robust inequalities are employed,
ESPPCC does not suffice in modeling the pricing problem.
A slight variation of the ESPPCC characterized by additional constraints
becomes mandatory to ensure correctness of the column generation approach.

The ESPPCC asks for the determination of the shortest path on a weighted network
between two vertices $s, t$
subject to a single global resource restriction characterized
by the amount of goods available for serving.
In the ESPPCC problem,
the elementarity condition imposes that no optimal path may
cover the same vertices two or more times.
The ESPPCC's underlying network may in the general case,
contain negative cost cycles.
The presence of negative cost cycles makes solving such problem NP-hard \parencite{dror1994}.
A dynamic programming algorithm
to tackle the elementary version of the problem
was proposed in \textcite{feillet2004}.
However, in the context of pricing, the problem has been traditionally relaxed into a SPPRC.
The SPPRC,
contrary to the elementarity version,
can be solved in pseudo-polynomial time with a much faster dynamic programming algorithm \parencite{desrochers1992}.
Unfortunately,
relaxing the elementarity condition leads to slower column generation and worsening of dual bounds,
see previous discussion in \cref{sec:intro-solving-the-pricing-problem}.

\medskip

In the case that the ESPPCC's weighted network contains no non-negative
cost cycles, the elementarity condition maybe safely dropped \parencite{beasley1989}.
A sufficient but not-necessary condition is when the reduced cost variables
are all positive: $\bar{c}_{ij} \ge 0 \quad \forall i, j \in V_0 \cup \Set*{s, t}$.
In such case, an optimal solution of the associated non-elementary SPPRC problem
is an optimal solution to the original elementary version of the problem \parencite{beasley1989}.
The associated SPPRC can be solved efficiently in pseudo-polynomial time.
Initial contributions for tackling the ESPPCC exploited this property.
The resource consumption constraint was relaxed through a Lagrangean relaxation
approach.
Through the usage of a branch-and-bound scheme and an off-the-shelf
standard shortest path algorithm, e.g. Dijkstra \parencite{sniedovich2006dijkstra},
they obtained feasible dual solutions over strictly positive weighted networks.
See the contributions of
\textcite{beasley1989, dumitrescu2003improved, carlyle2008, muhandiramge2009simultaneous},
just to name a few.
But as pointed out in \textcite{righini2004},
Lagrangian relaxation is effective only when the dualized variables
are positive for a good portion of the Lagrange multipliers search space,
which limit the effectiveness of these approaches in the context of pricing.

\subsection{Integer Programming Formulation}
\label{sec:espprc-integer-programming-formulation}

While \textcite{jepsen2008branchandcut} provides
an \textbf{undirected} symmetric network based formulation for the ESPPCC,
in our presentation we have preferred instead to re-adjust the more-general
\textbf{directed} symmetric network based ESPPRC formulation
provided in the works of \textcite{beasley1989, toth2002, toth2014}.
For this reason we need to readjust/extend the
mathematical notation that was originally provided for the undirected CVRP in \cref{sec:intro-cvrp-mathematical-notation}.

\medskip

The ESPPCC network is characterized by $N^\prime = N_0 + 2$ nodes,
where $N_0$, denotes the total number of customers in the original CVRP problem.
Two additional vertices $s = 0, t = N^\prime - 1$ are used to represent respectively the source and sink vertices.
These two together model a dissection of the depot node.
All the customers are edge-connected in both directions with one another.
Similarly, the source and sink vertex do not share direct connections together.

More formally, the ESPPCC is defined over a directed symmetric network
$G^\prime = \Tuple*{V^\prime, A^\prime}$,
where $V^\prime = V_0 \cup \Set*{s, t}$ denotes the set of vertices and $A^\prime$ the set of arcs.
The vertices $s = 0,\ t = N^\prime - 1$ denote respectively the source and sink versions of the depot node.
The vertex set $V^\prime$ has size $|V^\prime| = N^\prime = N_0 + 2$.
The edges set $A^\prime$ can be expressed as:
\begin{equation}
	\begin{aligned}
		A^\prime = \  & \Set*{\Tuple*{i, j} \mid i, j \in V_0,\ i \ne j} \\
		              & \cup \Set*{\Tuple*{s, i} \mid i \in V_0}         \\
		              & \cup \Set*{\Tuple*{i, t} \mid i \in V_0}.
	\end{aligned}
\end{equation}
The edge set $A^\prime$ has size $|A^\prime| = N^\prime(N^\prime - 1) + 2(N^\prime - 1)$.
In accordance to the CVRP problem,
we associate a resource consumption, or vertex demand,
to each node of the network: $q_i \quad \forall i \in V^\prime$.
The resource consumption semantics associated with each customer
$i \in V_0$ remains unaltered  from the CVRP, while for the source and sink vertices
we respectively fix $q_{s} = q_{t} = 0$.
We formally define $\deltaplus(S) = \Set*{(i, j) \in A^\prime \mid i \in S, j \notin S}$
to denote the out-arcs crossing the set $S \subset V^\prime$.
Likewise, we define $\deltaminus(S) = \Set*{(i, j) \in A^\prime \mid i \notin S, j \in S}$
to denote the in-arcs crossing the set $S \subset V^\prime$.
For brevity’s sake,
we also define $\deltaplus(i) = \deltaplus(\Set*{i})$, $\deltaminus(i) = \deltaminus(\Set*{i})$
to denote respectively the singleton version of the in and out arcs for vertices $i \in V^\prime$.
Note that the following conditions hold:
$\deltaplus(s) = V_0,\ \deltaminus(t) = V_0,\ \deltaminus(s) = \emptyset,\ \deltaplus(t) = \emptyset$.
We also define $A^\prime(S) = \Set*{\Tuple*{i, j} \in A^\prime \mid i, j \in S}$
to denote the set of arcs having both end points in set $S \subseteq V^\prime$.

As it was done in \cref{sec:column-generation-and-pricing-problem},
for each directed arc we associate a weight
$\bar{c}_{ij} \in \R \quad \forall \Tuple*{i, j} \in A^\prime$.
The weight $\bar{c}_{ij}$ represents the reduced cost of an arc $\Tuple*{i, j} \in A^\prime$.
Its definition is linked to the dual variables $\pi \in \R^N$ associated to
the RMP's constraints of \cref{eq:mp-K-routes,eq:mp-customers-visited-by-exactly-one-route}
through the following relationship:
\begin{equation}
	\bar{c}_{ij} = \begin{cases}
		c_{ij} - \frac{\pi_i + \pi_j}{2} & \quad \forall i, j \in V_0 \\
		c_{si} - \frac{\pi_0 + \pi_j}{2} & \quad \forall j \in V_0    \\
		c_{it} - \frac{\pi_i + \pi_0}{2} & \quad \forall i \in V_0    \\
		\infty                           & \quad \texttt{otherwise}
	\end{cases}
\end{equation}
Additional usage of robust inequalities (cuts or branching) manifest
additional negative term contributions, through additional dual variables,
directly to the reduced arc cost $\bar{c}_{ij}$.
See previous discussion in \cref{sec:intro-branching-and-cut-generation-within-bap-frameworks}.

A feasible path solution to the ESPPCC is a sequence $p = \Tuple*{p_0, p_1, \dots, p_s, p_{u+1}}$
with $p_0 = s, \ p_{u + 1} = t$
in which $\Set*{p_1, \dots, p_u} \subseteq V_0$ customers are visited.
Observe that there's no restriction in which customers need to be necessarily covered
by the path $p$.
Due to the elementarity condition, the path $p$ can cover a vertex at most once.
The path's resource consumption $q_p = q(p) = \sum_{j=0}^{u} q_{p_j}$
satisfies $q(p) \le Q$.
The optimal path $p^\star$ minimizes the overall path's reduced cost
$\bar{c}(p) = \sum_{i=0}^{u} \bar{c}_{p_i,p_{i+1}}$ across all
possible feasible choices for $p$.
Note that, regardless we have decided to model
the dual variables contributions $\pi \in \R^N$
at the arc level instead at the vertex level,
the path reduced cost still sums to:
$\bar{c}(p) = \sum_{i=0}^{u} c_{p_i,p_{i+1}} - \sum_{i=0}^{u} \pi_{p_i} = c(p) - \sum_{i=0}^{u} \pi_{p_i}$.
Therefore, $\pi_i \in R,\ i \in V_0$ can be interpreted as the
gained profit obtained in serving customer $i \in V_0$.
Similarly,
$\pi_0 \in R$ can be interpreted as an additional constant term
biasing the reduced cost of a feasible route $p$.

\medskip

We are now ready to provide a ESPPCC formal description through an Integer Programme (IP).
Let $x_{ij} \in \Set*{0, 1} \quad \Tuple*{i, j} \in A^\prime$ be a new set of binary variables:
$x_{ij} = 1$ if arc $\Tuple*{i, j} \in A^\prime$ is picked by the optimal path.
We can provide a formal mathematical formulation of the ESPPCC as an Integer Programme (IP):
\begin{align}
	\min_{x} \quad z_\mt{ESPPCC}(x) & =  \sum_{(i, j) \in A^\prime} \bar{c}_{ij} x_{ij}                                                     \label{eq:espprc-obj-function}                                                                \\
	                                & \sum_{i \in V_0} \frac{q_i}{2} \Expr*{ \ExprESPPIngoingEdges{i} + \ExprESPPOutgoingEdges{i} }  \le Q  \label{eq:espprc-resource-upper-bound-constraint}                                             \\
	                                & \ExprESPPIngoingEdges{i} = \ExprESPPOutgoingEdges{i}                                                  \qquad \forall i \in V_0          \label{eq:espprc-flow-conservation-constraint-customers}    \\
	                                & \ExprESPPOutgoingEdges[i]{s} = 1                                                                      \label{eq:espprc-source-emitting-flow}                                                        \\
	                                & \ExprESPPIngoingEdges[i]{t} = 1                                                                       \label{eq-espprc-sink-consuming-flow}                                                         \\
	                                & \ExprESPPEdgesWithin[S] \le |S| - 1                                                                   \qquad \forall S \subseteq V_0 \cup \Set*{s, t},\ |S| \ge 2 \label{eq:espprc-sec-constraints} \\
	                                & x_{ij}                   \in \lbrace 0, 1 \rbrace                                                     \qquad \forall \Tuple*{i, j} \in A^\prime    \label{eq:espprc-x-mip-var-bounds},
\end{align}
where \labelcref{eq:espprc-obj-function} is the objective function to be minimized, i.e. the overall reduced cost of the path.
Constraint \labelcref{eq:espprc-resource-upper-bound-constraint} is the resource consumption upper bound.
Constraint \labelcref{eq:espprc-flow-conservation-constraint-customers} imposes a flow conservation for all vertices except for the source and the sink.
Observe that \cref{eq:espprc-flow-conservation-constraint-customers} does not impose
restriction on whether customer $i \in V_0$ needs to be necessarily covered by the path.
Constraints \labelcref{eq:espprc-source-emitting-flow}, \labelcref{eq-espprc-sink-consuming-flow}
respectively impose an out-flow and in-flow of $1$ for the source and sink vertices.
Constraints \labelcref{eq:espprc-sec-constraints} are the Subtour Elimination Constraints (SEC) and
have a two-fold effect.
First, they avoid the formation of spurious tours in unconnected regions of the network, such
as the situation depicted in \cref{fig:espprc-example-with-spurious-subtours}.
Second, they impose the elementarity condition, restricting paths from visiting multiple times
the same vertices, such as the situation depicted in \cref{fig:espprc-example-non-elementary-solution}.
Constraint \labelcref{eq:espprc-x-mip-var-bounds} imposes integrality
and forces each arc $\Tuple*{i, j} \in A^\prime$ to be traversed at most once.

Notice that, due to the flow conservation constraint and to the bounds on the binary varibles of
\cref{eq:espprc-flow-conservation-constraint-customers,eq:espprc-x-mip-var-bounds},
we have that the following equality holds:
\begin{equation}
	\ExprESPPIngoingEdges{i} + \ExprESPPOutgoingEdges{i} = \begin{cases}
		2 & \texttt{if } i \in p \\
		0 & \texttt{otherwise}
	\end{cases}
	\quad \forall i \in V_0.
\end{equation}
Regardless of a constant factor of $2$,
the quantity $\ExprESPPIngoingEdges{i} + \ExprESPPOutgoingEdges{i}$
can be used to determine the number of times a  customer $i \in V_0$ is visited.

The resource consumption constraint of \Cref{eq:espprc-resource-upper-bound-constraint} is
expressed at the vertex level.
Equivalently, it is possible to express the same constraint by considering the resource consumption
at the arc level:
\begin{equation}
	\label{eq:espprc-resource-upper-bound-constraint-arc-level}
	\sum_{\Tuple*{i, j} \in A^\prime} \frac{q_i + q_j}{2} x_{ij} \le Q.
\end{equation}
Due to the structure of the optimal path $p^\star$ the total resource consumption
will not change regardless of whether
\labelcref{eq:espprc-resource-upper-bound-constraint}
or \labelcref{eq:espprc-resource-upper-bound-constraint-arc-level}
is used.

\begin{figure}[ht]
	\begin{tikzpicture}[>=stealth, main/.style = {draw, circle, fill=black!10!white}]
		\centering

		\node[main, ultra thick] (0) at (0.1\textwidth,-20bp) {0};
		\node[main] (1) at (0.25\textwidth,0bp){1};
		\node[main] (6) at (0.4\textwidth,+10bp){6};
		\node[main, ultra thick] (7) at (0.55\textwidth,+30bp) {7};

		\node[main] (2) at (0.7\textwidth, +20bp){2};
		\node[main] (3) at (0.8\textwidth, -0bp) {3};
		\node[main] (4) at (0.7\textwidth, -20bp) {4};
		\node[main] (5) at (0.6\textwidth, -0bp) {5};

		\draw [->,thick] (0) -- (1);
		\draw [->,thick] (1) -- (6);
		\draw [->,thick] (6) -- (7);

		\draw [->,thick] (2) -- (3);
		\draw [->,thick] (3) -- (4);
		\draw [->,thick] (4) -- (5);
		\draw [->,thick] (5) -- (2);
	\end{tikzpicture}

	\centering
	\caption{
		An example of a path containing spurious unconnected sub-tours,
		over a network with $V^\prime = \Set*{0, \dots, 7},\ s = 0,\ t = 7$.
		The SEC inequalities of \cref{eq:espprc-sec-constraints} prohibit the depicted situation from occurring.
	}
	\label{fig:espprc-example-with-spurious-subtours}
\end{figure}

\begin{figure}[ht]
	\begin{tikzpicture}[>=stealth, main/.style = {draw, circle, fill=black!10!white}]
		\centering

		\node[main, ultra thick] (0) at (0.1\textwidth,0bp) {0};
		\node[main] (1) at (0.3\textwidth,0bp){1};
		\node[main] (2) at (0.5\textwidth,0bp){2};
		\node[main] (3) at (0.6\textwidth,-30bp) {3};
		\node[main] (4) at (0.5\textwidth,-60bp) {4};
		\node[main] (5) at (0.4\textwidth,-30bp) {5};
		\node[main] (6) at (0.7\textwidth,0bp){6};
		\node[main, ultra thick] (7) at (0.9\textwidth,0bp) {7};

		\draw [->,thick] (0) -- (1);
		\draw [->,thick] (1) -- (2);
		\draw [->,thick] (2) -- (3);
		\draw [->,thick] (3) -- (4);
		\draw [->,thick] (4) -- (5);
		\draw [->,thick] (5) -- (2);
		\draw [->,thick] (2) -- (6);
		\draw [->,thick] (6) -- (7);
	\end{tikzpicture}

	\centering
	\caption{
		An example of a path not satisfying the elementarity constraints,
		over a network with $V^\prime = \Set*{0, \dots, 7},\ s = 0,\ t = 7$.
		The SEC inequalities of \cref{eq:espprc-sec-constraints} prohibit the depicted situation from occurring.
	}
	\label{fig:espprc-example-non-elementary-solution}
\end{figure}

As it is done in \textcite{beasley1989},
if one wishes to avoid the formation of spurious unconnected subtours
while relaxing the elementarity condition,
it can achieve so by substituting \cref{eq:espprc-sec-constraints} in favor
of a big-M constraint of the form:
\begin{equation}
	\label{eq:espprc-relaxed-elementarity}
	\ExprESPPEdgesWithin[S] \le M \ExprESPPEdgesCrossing[S] \quad \forall S \subseteq V_0 \cup \Set*{s}
\end{equation}
where $M \in \R_+$ denotes an arbitrary large positive constant.
\Cref{eq:espprc-relaxed-elementarity} avoids the situation depicted
in \cref{fig:espprc-example-with-spurious-subtours}
while permitting the situation depicted in \cref{fig:espprc-example-non-elementary-solution}.
Modeling the pricing as an SPPCC, can greatly reduce
the column generation running time, at the cost
of generating significantly weaker dual bound improvements.
Refer back to the discussion in \cref{sec:intro-solving-the-pricing-problem}
for more details.

Additional valid inequalities and a branch-and-cut framework
for the ESPPCC was provided in the work of \textcite{jepsen2008branchandcut}.
In this thesis we are mainly concerned with the CPTP in the context of pricing.
Therefore we will not spend more efforts in discussing the ESPPCC.

\section{The Capacitated Profitable Tour Problem}
\label{sec:the-capacitated-profitable-tour-problem}

The Capacitated Profitable Tour Problem, abbreviated as CPTP,
belongs to the group of the so-called
Travelling Salesman Problems with Profits (TSPP) (see \cite{feillet2005}),
and therefore presents many similarities with other
studied optimization problems such as:
(i) the Orienteering Problem (OP) \parencite{golden1987, laporte1990}
(ii) the Profitable Tour Problem (PTP) \parencite{dellamico1995},
(iii) the Prize Collecting Traveling Salesman Problem (PCTSP) \parencite{balas1989prize, balas1995prize},
(iv) the Capacitated m-Ring-Star Problem (CmRSP) \parencite{baldacci2007capacitated}.
Refer to \textcite{letchford2013} for the Steiner extension of these problems.

The OP, an NP-hard problem \parencite{laporte1990}, asks for a route serving
a subset of customers with associated profits.
In OP, the route length is bounded in terms of number of vertices visited,
whereas in CPTP the route is bounded in terms of the accumulated demand served.
Using the reduction to Hamiltonian circuit problem
employed to prove that the OP is NP-hard \parencite{laporte1990},
the same reasoning can be applied to conclude that CPTP is also NP-hard.
The OP may be regarded as a special case of an edge-capacitated CPTP with unit demands.

Refer to \textcite{fischetti1998, gendreau1998} for contributions
exposing BAC frameworks and valid inequalities for the OP.
Many of such inequalities (but not all) are also valid for the CPTP \parencite{jepsen2014}.
We will introduce the CPTP in the context of pricing for the CVRP.
We will study the basic version of the CPTP,
i.e. we will not consider additional non-robust inequalities
at the RMP level that could change the structure of the problem.

\medskip

In more detail,
the CPTP is a combinatorial optimization delivery problem,
in which,
given as input a set of customers
defined over a fully connected undirected network, the goal resides in finding
a resource constrained elementary tour (or circuit) starting from a common point called the depot,
that minimizes the overall travel distance while maximizing
the profits associated in serving only a subset of the customers.
The profit value is gained only when serving occurs.
The elementarity constraint dictates that no vertex
can be visited twice or more, thus implying that the earnable profit
per customer is available only once.

The CPTP shows up naturally as a sub-problem
in the column generation approach employed in solving the CVRP.
It is an important special case of the
ESPPCC when the source and sink vertices encode the same node (the depot node).
In the traditional CPTP formulation \parencite{jepsen2014},
the profit is encoded at the node level
instead of being included directly in the definition
of the reduced cost variable $\bar{c}_{ij} \quad \forall i, j \in V_0 \cup \Set*{0}$
as it is done in the ESPPCC.
Contrary to the formulation of the more general ESPPRC,
the CPTP is much easier to describe.
However, due to its narrower scope,
it is a problem which is far less addressed in the literature.
To the best of our knowledge, the few contributions studying the CPTP
in the context of pricing for the CVRP can be found in \textcite{bixby1999, jepsen2011,jepsen2014}.

The presentation of this section,
along with the implementation discussed in \cref{sec:implementation-chapter}
owes a lot to the original efforts of \textcite{jepsen2014},
where they laid the foundational framework for the CPTP.

\subsection{Integer Programming formulation}
\label{sec:cptp-integer-programming-formulation}

We provide a CPTP formulation, which closely follows the formulation provided in \textcite{jepsen2014}.
In terms of mathematical notation, we will use and extend the original
mathematical constructs introduced in \cref{sec:intro-cvrp-mathematical-notation} for the CVRP.

Let $p_i \in \R,\ p \ge 0$ denote the profit associated in visiting a vertex $i \in V$.
The profit function is directly linked to the dual variables $\pi \in \R^N$ associated to
the RMP's constraints of \cref{eq:mp-K-routes,eq:mp-customers-visited-by-exactly-one-route}.
Namely, $p_i = \pi_i \quad \forall i \in V$.

Let $x_{e}$, $y_{i} \in \Set*{0, 1}$ be two sets of binary variables which respectively
model whether an undirected edge $e \in E$ or vertex $i \in V$ is covered by the optimal CPTP route.
We can provide a formal mathematical description of the CPTP through an Integer Programme (IP) formulation:
\begin{align}
	\min_{x,y} \quad z_\mt{CPTP}(x, y) & = \ExprCptpObjValDef                     & \label{eq:cptp-obj-function}                                                                          \\
	                                   & y_0 = 1                                  & \label{eq:cptp-depot-part-of-tour-constraint}                                                         \\
	                                   & \ExprCptpDemandSum  \le Q                & \label{eq:cptp-resource-upper-bound-constraint}                                                       \\
	                                   & \ExprCptpEdgesIncident{i}  = 2 y_i       & \quad \forall i \in V         \label{eq:cptp-flow-conservation-constraint}                            \\
	                                   & \ExprCptpFlowExiting{S} \ge 2 y_{i}      & \quad \forall i \in S,\ \forall S \subseteq V_0,\ \SetSize*{S} \ge 2 \label{eq:cptp-gsec-constraints} \\
	                                   & x_{e}                   \in \Set*{0, 1}  & \quad \forall e \in E               \label{eq:cptp-x-mip-var-bounds}                                  \\
	                                   & y_{i}                    \in \Set*{0, 1} & \quad \forall i \in V             \label{eq:cptp-y-mip-var-bounds},
\end{align}
where \labelcref{eq:cptp-obj-function} is the objective function to be minimized, i.e. the cost
accounting for the total travel distance and the profit in serving a subset of the vertices.
Constraint \labelcref{eq:cptp-depot-part-of-tour-constraint} ensures that the depot is always covered by the optimal route.
Constraint \labelcref{eq:cptp-resource-upper-bound-constraint} is the resource consumption upper bound.
Constraint \labelcref{eq:cptp-flow-conservation-constraint} has a twofold effect.
First, it functions as a form of directed flow conservation.
Second, it binds the $x,\ y$ sets of variables together;
namely, it ensures that the if a vertex $i \in V$ is covered by an optimal route ($y_i = 1$)
then the number of incident edges in that node sums to $2$, otherwise to $0$.
Constraint \labelcref{eq:cptp-gsec-constraints}
are the so-called Generalized Subtour Elimination Constraints (GSECs),
which avoid the formation of spurious tours in unconnected regions of the network.
There is an exponential number of GSECs,
therefore for performance reasons,
these constraints cannot be inserted statically in a MIP solver,
and must instead be separated exactly at least for integral solutions.
Constraints \labelcref{eq:cptp-x-mip-var-bounds,eq:cptp-y-mip-var-bounds} respectively
impose bounds and integrality conditions
on the number of traversal for edges $e \in E$ and vertices $i \in V$.
The IP model consists of $\frac{N^2 + N}{2}$ number of binary variables and an exponential number of constraints.
If we ignore the GSECs and variable bounds, the IP model owns a total of $N + 2$ constraints.
Constraints \labelcref{eq:cptp-gsec-constraints}
can also be rewritten in a different form \parencite{wolsey2020integer}:
\begin{equation}
	\label{eq:cptp-gsec-constraints-v2}
	\sum\limits_{e \in E(S)} x_e \le \sum\limits_{i \in S \setminus \Set*{j}} y_i \quad \forall j \in S,\ \forall S \subseteq V_0,\ |S| \ge 2.
\end{equation}
Depending on the size of the subset $S \subseteq V_0, S \ge 2$,
\cref{eq:cptp-gsec-constraints-v2} may be sparser than \cref{eq:cptp-gsec-constraints}.
Equivalently, the constraint of \cref{eq:cptp-resource-upper-bound-constraint}
can be defined at the edge level:
\begin{equation}
	\label{eq:cptp-resource-upper-bound-constraint-edge-level}
	\sum_{e = \Set*{i, j} \in E} \frac{q_i + q_j}{2} x_e \le Q.
\end{equation}

Note that model
\labelcref{eq:cptp-obj-function,eq:cptp-depot-part-of-tour-constraint,eq:cptp-resource-upper-bound-constraint,eq:cptp-flow-conservation-constraint,eq:cptp-gsec-constraints,eq:cptp-x-mip-var-bounds,eq:cptp-y-mip-var-bounds}
does not permit for single-customer routes.
Instead of modifying the formulation to allow for this edge case,
it may be simpler to leave the model unchanged.
By employing a brute-force algorithm in $\Theta(N)$ time we can scan for improving single-customer solutions.
After the resolution of the IP model,
we check for improving single-customer solutions and,
if any exist,
we update the incumbent as appropriate.

\section{Additional Valid Inequalities}
\label{sec:cptp-additional-valid-inequalities}

In this section we present additional valid inequalities for the CPTP problem.
While these additional inequalities are not strictly necessary
to ensure the correctness of the solution approach,
when embedded inside an efficient branch-and-cut framework,
through efficient separation techniques,
can massively strengthen the linear relaxation and thus speed up the resolution process.

The inequalities that we present follow closely the presentation provided in \textcite{jepsen2014}.
Refer to the \textcite{jepsen2014} for a more complete set of valid inequalities.
In this thesis we will concentrate mainly on two additional valid inequalities:
the Rounded Capacity Constraints (RCC) discussed in \cref{sec:cptp-rcc},
and the Generalized Large Multistars (GLM) discussed in \cref{sec:cptp-glm}.
We have concentrated our attentions on only these two sets of inequalities
for two main reasons.
First, the work of \textcite{jepsen2014} showed that these two inequalities are the most frequently
separated, suggesting that their inclusion would likely speed up the resolution process.
Second,
they can be effectively separated through the same single procedure
shared also with the GSECs inequalities separation.
Thus reducing the implementation effort required for their inclusion.
Refer to the implementation details of \cref{sec:impl-separation-techniques}.

\subsection{Rounded Capacity Constraints (RCC)}
\label{sec:cptp-rcc}

The Rounded Capacity Constraints, RCC for short, were first presented
for the CVRP  in \textcite{laporte1983}
and were already presented in \cref{sec:intro-cvrp-two-index-flow-formulation}.

Recall that $q(S) = \sum_{i \in S} q_i$ represents
the total demand in serving all vertices in set $S \subseteq V$.
The RCC constraints function as a capacity constraint
by imposing that any customer set $S \subseteq V_0$ is crossed by a number of edges
not smaller by the number of trucks that would be necessary to serve all customers in $S \subseteq V_0$.
The RCC can be extended to the CPTP problem \parencite{jepsen2014}:
\begin{equation}
	\label{eq:cptp-rcc-inequality-original}
	\ExprCptpFlowExiting{S} \ge 2 \ceil*{\frac{\sum\limits_{i \in S} q_i y_i}{Q}} \quad \forall S \subseteq V_0,\ |S| \ge 1.
\end{equation}
\Cref{eq:cptp-rcc-inequality-original}
is very similar to the CVRP's RCC of \cref{eq:two-index-flow-ccc-bpp-lb},
except that the right-hand side is composed of model's variables instead of constants.
Notice that such approach is correct since $\sum_{e \in \delta(S)} x_e \in \Z_+$.
Unfortunately, the ceil function is a non-linear operation and therefore
such constraint cannot be inserted in most MIP solvers.

Fortunately, \textcite{baldacci2007capacitated} provides a way to
combat this undesirable aspect at the cost of obtaining a looser bound of the RCC.
Namely, given $\alpha, \beta, \gamma \in Z_+$, with $\alpha > \gamma$ and
$\fmod{\alpha}{\gamma} \ne 0$, the following result holds:
\begin{equation}
	\label{eq:baldacci-rcc-theorem-weaker-bound}
	\ceil*{\frac{\alpha - \beta}{\gamma}} \ge \ceil*{\frac{\alpha}{\gamma}} - \frac{\beta}{\fmod{\alpha}{\gamma}}.
\end{equation}
\textcite{jepsen2014} claims that \cref{eq:baldacci-rcc-theorem-weaker-bound}
holds even when $\alpha > \gamma$ is not satisfied
(unfortunately the authors don't provide further explanations on why this is the case).

By picking $\alpha = q(S) = \sum_{i \in S} q_i,\ \beta = \sum_{i \in S} q_i\Expr*{1 - y_i},\ \gamma = Q$,
we obtain the following linear, but weaker, RCC:
\begin{equation}
	\label{eq:cptp-rcc-inequality}
	\sum\limits_{e \in \delta(S)} x_e \ge 2 \Expr*{\ceil*{\frac{q(S)}{Q}} - \frac{\sum\limits_{i \in S} q_i \Expr*{1 - y_i}}{\ExprQr(S)}} \quad \forall S \subseteq V_0,\ |S| \ge 1,
\end{equation}
where $\ExprQr(S) = \fmod{q(S)}{Q}$ is the remainder capacity associated
to an arbitrary set $S \subseteq V$.
Notice, that \cref{eq:baldacci-rcc-theorem-weaker-bound} guarantees correctness
of the \cref{eq:cptp-rcc-inequality} only if $q_i \in Z_+ \quad \forall i \in V_0$ holds.
Fortunately this is the case for most, if not all, routing problems considered in the literature.

As \textcite{jepsen2014} points out,
no exact separation procedure is currently known for the RCC inequalities,
for this reason,
they suggest using the sets identified by the separation of the multistar inequalities,
such as the GLM separation procedure.
See the next \cref{sec:cptp-glm}.

\subsection{Generalized Large Multistar (GLM)}
\label{sec:cptp-glm}

The Generalized Large Multistar inequalities,
GLM for short,
were first proposed in \textcite{gouveia1995} for the CVRP.
In \textcite{letchford2002}
the GLM were further generalized to the so-called Knapsack Large Multistar (KLM) inequalities.
\textcite{letchford2006} study the effectiveness of the multistar family of cuts
when applied to the vehicle routing problems.
The multistar cuts represent a set of inequalities that are related to
the intersection between a 0--1 knapsack problem and the node-capacitated profitable tour problem (CPTP).

Consider the capacity inequality:
\begin{equation}
	\label{eq:cptp-weaker-rcc}
	\sum_{e \in \delta(S)} x_e \ge \frac{2}{Q} \sum_{i \in S} d_i y_i \quad \forall S \subseteq V_0,\ |S| \ge 2,
\end{equation}
which is a weaker form of the RCC in \cref{eq:cptp-rcc-inequality-original},
where ceiling is not performed.
\Cref{eq:cptp-weaker-rcc} can be improved to obtain the GLM inequalities,
by noting that  nodes $j \notin S$ are also visited when the crossing edges $\delta(S)$ are used.

The GLM inequalities can be expressed as \parencite{jepsen2014}:
\begin{equation}
	\label{eq:cptp-glm-inequality}
	\sum\limits_{e \in \delta(S)} x_e \ge \frac{2}{Q} \Expr*{
		\sum\limits_{i \in S} q_i y_i +
		\sum\limits_{\EqStackTwo{e = \Set*{i, j} \in \delta(S)}{i \in S,\ j \notin S}} q_j x_e
	}
	\quad \forall S \subseteq V_0,\ |S| \ge 2.
\end{equation}

The general idea behind multistar inequalities is the following.
The vehicle must have sufficient capacity to serve
all the customers covered by $S$
and all other customers outside $S$
which are connected to nodes in $S$ through a crossing edge $e = \Set*{i, j},\ i \in S,\ j \notin S$.
The "capacity" in GLM is measured in terms of the number of vehicles,
which is obtained from dividing by $Q$ the demand of the set $S$ plus
any customer visited directly from the set $S$ itself.
GLM inequalities are efficiently separable with
an exact polynomial time procedure by finding minimum
$(u, v)$-cuts $\forall u \in V_0,\ v = 0$
in a directed flow-network with capacities which depend on the vertex $u$
\parencite{letchford2006, jepsen2014}.
A simpler approach consists instead of using a heuristic approach
by minimizing only the term $\sum_{e \in \delta(S)} x_e$ as to stress the violation.
This can be achieved by computing $(u, v)$-cuts $\forall u \in V_0,\ v = 0$
in a directed flow network where capacities are given exclusively from the
fractional solution $x^\star _{\Set*{i, j}}\quad \forall i, j \in V$.
